---
title: "Accelerometer Prediction"
author: "Dat Laffy Taffy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary
Data was collected from accelerometers on the belt, forearm, arm, and bar of 6 participants. They were asked to perform lifts correctly and incorrectly in 5 different ways. 

The purpose of this report was to analyse the data and build a model using machine learning methods, to predict the type of lift performed.

A Random Forest prediction model was found to be 99.42% accurate, with an appropriate out of sample error.


# Data Processing

```{r}
# Load necessary packages
library(data.table)
library(dplyr)
library(caret)
library(randomForest)

# Download the data
url_training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
training_file <- "pml-training.csv"
download.file(url_training, destfile = training_file, mode = "wb")
training <- read.csv("pml-training.csv")

url_testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
testing_file <- "pml-testing.csv"
download.file(url_testing, destfile = testing_file, mode = "wb")
testing <- read.csv("pml-testing.csv")

# Clean the data
# Removing Variables with Nearly Zero Variance
NZV <- nearZeroVar(training)
training <- training[, -NZV]
# Removing variables that are mostly NA
AllNA <- sapply(training, function(x) mean(is.na(x))) > 0.95
training <- training[, AllNA==FALSE]
# Removing columns that are not related to the predictors
training <- training[, -(1:7)]
# Converting the classe variable to a factor 
training$classe <- as.factor(training$classe)
```


# Prediction Models

**Split the training data into 70% training and 30% validation datasets**

```{r}
set.seed(1234) 
partition <- createDataPartition(y = training$classe,
                                   p = 0.7,
                                   list = FALSE)
data_training <- training[partition,]
data_validation <- training[-partition,]
```


**Build a prediction model - Random Forest**

*(Note: This takes a few minutes so be patient)*

```{r}
set.seed(1234)
rfModel <- randomForest(classe ~ ., data = data_training, ntree = 100)
rfPrediction <- predict(rfModel, data_validation)
confusionMatrix(data_validation$classe, rfPrediction)
```

*Observation:* This model results in 99.42% accuracy. It is a suitable model for prediction.


**Cross validation & Expected out of sample error**

```{r}
oob_error <- 1 - rfModel$err.rate[nrow(rfModel$err.rate), "OOB"]
cat("Out-of-Bag (OOB) Error:", oob_error, "\n")
```

*Observation:* The out of sample error rate is good. It is a suitable model for prediction.


# 20 test cases

```{r}
set.seed(1234)
predict(rfModel, testing)
```